\documentclass{article}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\renewcommand{\thesubsection}{\arabic{subsection}}
\definecolor{codebg}{RGB}{248,248,248}
\definecolor{keywordcolor}{RGB}{0,0,180}
\definecolor{stringcolor}{RGB}{163,21,21}
\definecolor{commentcolor}{RGB}{0,128,0}
\definecolor{numbercolor}{RGB}{128,0,128}
\definecolor{decoratorcolor}{RGB}{200,100,0}

\lstdefinestyle{py}{
  language=Python,
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\small,
  keywordstyle=\color{keywordcolor}\bfseries,
  stringstyle=\color{stringcolor},
  commentstyle=\color{commentcolor}\itshape,
  numberstyle=\tiny\color{gray},
  numbers=left,
  stepnumber=1,
  numbersep=8pt,
  showstringspaces=false,
  breaklines=true,
  frame=single,
  rulecolor=\color{black!30},
  tabsize=2,
  morekeywords={self},
}
\hypersetup{
            pdfauthor={Xavier Sebastian A}, 
            pdftitle={Compsci:790-HW3}, 
            pdfsubject={HW3}, 
            pdfkeywords={GPU,Triton,JAX,Pallas,Kernal Fusion,JIT},
            pdfproducer={LaTeX with hyperref},
            pdfcreator={pdflatex},
            pdfborder={0 0 0}, 
            pdfstartview=FitH
        }

\title{Homework 3}
\author{Xavier Sebastian A}
%\date{\today}

\begin{document}

\maketitle
\section{Conceptual Questions}
\subsection{Explain the difference between tracing-based and scripting-
based JIT compilation. Give a concrete example of a Python function
that would work correctly with scripting but fail with tracing}
\subsubsection*{Answer:}
Tracing-based JIT runs the function once and records the operations that happen during that execution. 
It builds a computation graph based on that single run. 
Because of this, it cannot correctly handle control flow that depends on input values.

Scripting-based JIT analyses the program code directly and converts it into an intermediate representation. 
Since it understands the structure of the code, it can correctly handle conditions, loops, and other control flow statements.
The following example demonstrates scripting-based compilation in PyTorch \cite{jit_compilation_ml}:

\begin{lstlisting}[style=py]
@torch.jit.script # Example decorator
def conditional_op(x, threshold):
  # Scripting compiler parses this structure
  if x.sum() > threshold:
    result = x * 2
  else:
    result = x + 1
  return result
\end{lstlisting}
Tracing may record only one branch of the if statement depending on the input used during tracing. 
If the condition later changes, the result may become incorrect. 
Scripting works correctly because it keeps both branches in the compiled graph.

\subsection{Consider the following JAX function:}
\begin{lstlisting}[style=py]
@jax.jit
def process_batch(data, batch_size):
    num_batches = data.shape[0] // batch_size
    results = []
    for i in range(num_batches):
        batch = data[i * batch_size:(i + 1) * batch_size]
        results.append(jnp.sum(batch ** 2))
    return jnp.array(results)
\end{lstlisting}
What potential performance issue might arise with this JIT-compiled function? How would you fix it?
\subsubsection*{Answer:}
The function creates a Python list and appends values inside a loop. 
Python loops and list operations reduce the efficiency of JAX because they prevent proper optimisation during compilation.
JAX works best when operations are written in a vectorised form using array operations. 
A better solution is to reshape the data and compute the result using built in array functions.
Fixed version:
\begin{lstlisting}[style=py]
@jax.jit
def process_batch(data, batch_size):
    num_batches = data.shape[0] // batch_size
    data = data[:num_batches * batch_size]
    data = data.reshape(num_batches, batch_size)
    return jnp.sum(data ** 2, axis=1)
\end{lstlisting}
This removes the Python loop and allows JAX to perform faster compiled execution.

\subsection{The lecture mentioned that operator fusion can reduce memory traffic. 
For the expression $y = \sin(\cos(x)) + \exp(x)$ calculate:}

\begin{enumerate}
\item[(a)] The number of GPU kernel launches and global memory
reads/writes in eager mode (assume each operation reads inputs from
global memory and writes outputs to global memory).

\item[(b)] The number in fused mode (one kernel, intermediate values
stored in registers).

\item[(c)] The theoretical memory bandwidth reduction factor.
\end{enumerate}

\subsubsection*{Answer:}
\begin{enumerate}
    \item [(a)]
    Each operation runs as a separate GPU kernel.
    Operations performed:
    \begin{itemize}
    \item $\cos(x)$
    \item $\sin(\cdot)$
    \item $\exp(x)$
    \item addition
    \end{itemize}
    Total kernel launches $= 4$.Each operation reads data from global memory and writes the results back to global memory.

\item[(b)]
All operations are combined into a single GPU kernel.Total kernel launches $= 1$.Intermediate values remain inside GPU registers instead of being written to global memory.

\item[(c)]
In eager execution there are roughly $8$ memory operations ($4$ reads and $4$ writes).In fused execution there are $2$ memory operations ($1$ read and $1$ write).
Therefore, the theoretical reduction factor is $\frac{8}{2} = 4$.So the memory traffic is reduced by approximately four times.
\end{enumerate}

\subsection{Why does dynamic control flow (e.g., if statements depending
on tensor values) pose a challenge for JIT compilation? Explain both the
JAX and PyTorch approaches to handling this.}


\subsubsection*{Answer:}

Dynamic control flow occurs when program execution depends on values that are known only at runtime. JIT compilers typically construct a static computation graph before execution. If the program contains branches that depend on runtime tensor values, the compiler cannot determine the correct execution path during compilation.

JAX addresses this issue by providing specialised control-flow primitives such as \texttt{lax.cond}, \texttt{lax.scan}, and \texttt{lax.while\_loop}. These primitives explicitly represent control flow within the compiled graph.

PyTorch addresses the issue using \texttt{TorchDynamo}, which traces Python execution and captures graph segments dynamically. When unsupported control flow appears, the system may partially compile the program or fall back to eager execution.

\subsection{When you run \texttt{torch.compile} on a function for the first time,
it’s slower than the eager version. Why? What specific steps happen
during this first call that don’t happen in subsequent calls?}

\subsubsection*{Answer:}

The first execution of a function compiled with \texttt{torch.compile} involves several additional steps before the program runs.

During the first call:

\begin{itemize}
\item \texttt{TorchDynamo} traces the Python function and extracts a computation graph.
\item The graph is analysed and optimised.
\item \texttt{TorchInductor} or another backend compiler generates optimised kernels.
\item The generated code is compiled for the target hardware.
\item The compiled graph is cached for reuse.
\end{itemize}

These compilation steps introduce overhead during the first execution. Later executions reuse the cached compiled graph and therefore run significantly faster.


\begin{thebibliography}{9}

\bibitem{jit_compilation_ml}
JIT Tracing vs Scripting
\textit{Tracing vs. Scripting Approaches.}  
\url{https://apxml.com/courses/compiler-runtime-optimization-ml/chapter-7-jit-compilation-ml/jit-tracing-vs-scripting}



\end{thebibliography}
\end{document}

